---
title: "Chicklet boxplots with {ggchicklet2}"
output:
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Chicklet boxplots with {ggchicklet2}}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "## ",
  fig.width = 6,
  fig.height = 4,
  out.width = "100%"
)
```

`geom_chicklet_boxplot()` is a drop-in replacement for
`ggplot2::geom_boxplot()` whose box body is rendered as a rounded
rectangle. It is implemented as a real `ggproto` (`GeomChickletBoxplot`)
that inherits from `GeomBoxplot` and uses `StatBoxplot` for stat
computation, so:

- it supports the full boxplot aesthetic set (`x`/`y`, `lower`, `upper`,
  `middle`, `ymin`, `ymax`, `fill`, `colour`, `linewidth`, `linetype`,
  `alpha`, `weight`);
- it understands `orientation = "x"`/`"y"` and `coord_flip()`;
- it dodges correctly when you map a secondary aesthetic such as `fill`
  to a second variable; and
- it accepts the modern boxplot customization arguments (`outlier.*`,
  `whisker.*`, `staple.*`, `median.*`, `box.*`, plus the `*_gp` lists)
  that ggplot2 ≥ 3.5.0 exposes.

```{r setup}
library(ggplot2)
library(ggchicklet2)
```

## Single grouping

```{r single-group}
ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_chicklet_boxplot(
    radius      = grid::unit(4, "pt"),
    staplewidth = 0.5
  ) +
  theme_minimal()
```

## Dodged: secondary grouping via `fill`

This is the case where the original list-of-layers prototype broke down —
because the inner `dodge` was based only on the x variable. The real
`ggproto` version picks up the `fill` grouping automatically.

```{r dodged}
mt <- transform(mtcars,
                cyl  = factor(cyl),
                gear = factor(gear))

ggplot(mt, aes(cyl, mpg, fill = gear)) +
  geom_chicklet_boxplot(
    radius      = grid::unit(3, "pt"),
    staplewidth = 0.5
  ) +
  theme_minimal()
```

## Horizontal orientation

Just swap the aesthetics — the underlying `flipped_aes` machinery is
inherited from `GeomBoxplot`.

```{r horizontal}
ggplot(iris, aes(Sepal.Length, Species, fill = Species)) +
  geom_chicklet_boxplot(radius = grid::unit(6, "pt")) +
  theme_minimal()
```

## Pairs nicely with `ggsignif`

Because `geom_chicklet_boxplot()` is a proper geom whose layer data
includes the full set of boxplot summary columns, downstream packages
that introspect ggplot layer data (such as `ggsignif`, `ggpubr`,
`ggrepel`) work without modification:

```{r ggsignif, eval = requireNamespace("ggsignif", quietly = TRUE)}
library(ggsignif)

ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_chicklet_boxplot(radius = grid::unit(4, "pt"),
                        staplewidth = 0.5) +
  geom_signif(
    comparisons = list(c("setosa", "versicolor"),
                       c("versicolor", "virginica")),
    map_signif_level = TRUE,
    step_increase = 0.08
  ) +
  theme_minimal()
```

## Caveats

- `notch = TRUE` is intentionally unsupported; rounded corners and notch
  geometry are incompatible. The argument is accepted for signature
  compatibility but is silently ignored (and a warning is emitted if you
  set it explicitly).
- The legend key is currently inherited from `GeomBoxplot` rather than
  being a rounded mini-boxplot. This is a known cosmetic limitation.
